---
title: "CDCS2024NHT-teaching"
author: "Fang Yang"
date: "2024-05-09"
output: bookdown::pdf_document2
toc: false
---

# Preparation

```{r results='hide', warning=FALSE, message=FALSE}
# Load packages
library(tidyverse)
library(patchwork)
library(kableExtra)
``` 

```{r results='hide', warning=FALSE, message=FALSE}
# Read data 
data <- read_csv("Instadata.csv")
```

# Data Wrangling

```{r}
# Inspect data
dim(data)

glimpse(data)
summary(data)


data$Group <- as.factor(data$Group)
levels(data$Group)
```

# Descriptive Statistics

## Contingency Table

```{r}
tbl_stats <- data %>%
  group_by(Group) %>% 
  summarise(n = n(),
            M = mean(Time),
            SD = sd(Time),
            Min = min(Time),
            Max = max(Time))
tbl_stats
```
 
### Think point

Any problems?

How to fix them?

### Missing values 

```{r}
data <- data %>% drop_na(Time)

# check the data
tbl_stats <- data %>%
  group_by(Group) %>% 
  summarise(n = n(),
            M = mean(Time),
            SD = sd(Time),
            Min = min(Time),
            Max = max(Time))
tbl_stats
```


### Impossible values 


```{r}

isTRUE(data$Time > 0)

data <- data %>%
  filter(Time > 0)

# check the data
tbl_stats <- data %>%
  group_by(Group) %>% 
  summarise(n = n(),
            M = mean(Time),
            SD = sd(Time),
            Min = min(Time),
            Max = max(Time))
tbl_stats
```


## Visualisation

```{r fig.show='hide', message=FALSE, warning=FALSE}
plt_hist <- ggplot(data, aes(x = Time, after_stat(density))) +
  geom_histogram(fill = "darkgreen") +
  geom_density() +
  facet_wrap(~Group) +
  labs(x = "Time spent on Instagram (min)", 
       y = "Density",
       title = "(a)")
plt_hist

plt_box <- data %>%
  ggplot(aes(x=Group, y = Time, fill=Group)) +
  geom_boxplot() +
  labs(x = "Group", 
    y = "Time spent on Instagram daily(min)",
       title = "(b)") +
  theme() 
plt_box

```

```{r message=FALSE, warning=FALSE}
plt_hist | plt_box
```


# Null Hypothesis Testing

Our Alternative Hypothesis of the research is that university students on 
average spend more time on Instagram daily then full-time exployees.

The Null Hypothesis is there is no different between the two groups.

$H0$ : mu = 0
$H1$ : mu > 0

Because our hypothesis is one-tailed, i.e., has a direction. we need to 
correctly specify the "alternative = " parameter. We will firstly check the 
reference group, and then specify the "alternative = " accordingly.

```{r}
levels(data$Group)
```

Our hypothesised direction is Uni student > FT employee. 
Given that our reference level is FT employee, we need to specify alternative as
"less". This tell r that the alternative hypothesis assumes the group at the 
reference level (FT employee) is smaller than the group at the critical level
(Uni student). 

To make things easier, we relevel the factor so that University students are the 
reference group. We can then specify "alternative = " to "greater".

```{r}
data$Group <- relevel(data$Group, ref= "Unistudent")
```

## Two-sample t-test

Next we perform a two-sample t-test. We use the default significance level .05.

```{r}
t_test <- t.test(data$Time ~ data$Group, mu = 0, alternative = "greater", 
                 var.equal = TRUE)

t_test
```

### Interpretation

At the 5% significance level, a two-sample t-test was conducted to investigate
the amount of time spent on Instagram by university students and full-time 
employees. Results show that university students on average spend significantly 
more time (mean = 49.70 minutes) than than full-time employees (mean = 36.95 
minutes) on Instagram everyday (t(df=150)= 15.10, p < .001). 

## Confidence Intervals

We have got the results of our t-test. But how confident are we about our
results? To answer this question, we need to calculate the confidence intervals.

```{r}
t_test_CI <- t.test(data$Time ~ data$Group, mu = 0, alternative = "two.sided", 
                 var.equal = TRUE) 
t_test_CI 
```

### Interpretation 
 
We are 95% confident that university students on average spend between 11.08 and
14.42 more minutes on Instagram than full-time employees.  

### Think Point 

What if we want to use a more rigid significance level (e.g., alpha = .01)?

Tip: you just need to specify conf.level = .99.

# Exercise: YOUR TURN

Perform a two-sample t-test at significance level of .01 to answer the
research question.

```{r}
# Perform a t-test at alpha = .01

t_test_2 <- t.test(data$Time ~ data$Group, mu = 0, alternative = "less", 
                  conf.level=.99,
                 var.equal = TRUE)

t_test_2
```


```{r}
# get confidence intervals

t_test_2_CI <- t.test(data$Time ~ data$Group, mu = 0, alternative = "two.sided", 
                  conf.level=.99,
                 var.equal = TRUE)

t_test_2_CI
```

How would you interpret the results?
 